---
title: 'NES-LTER: Comparison between CTD and sampled chlorophyll concentration estimates'
output:
  html_document:
    df_print: paged
---

This notebook is an R translation of Joe Futrelle's python notebook that combines chlorophyll concentration estimates derived from a CTD-mounted fluorometer with corresponding estimates derived from lab processing of samples. This enables confirming that the estimates match, which aids in the decision of when to take samples.

First, read chlorophyll and CTD data from the API

```{r}
library(tidyverse)

BASE_URL = 'https://nes-lter-data.whoi.edu/api/'
chl = read_csv(paste0(BASE_URL, 'chl/en608.csv'))
btl = read_csv(paste0(BASE_URL, 'ctd/en608/bottles.csv'))
```

Take a look at the first few rows

```{r}
head(btl)
```

Samples are replicated, so the chlorophyll concentration needs to be averaged across those replicates. This uses a function called average_replicates.

Because dplyr uses non-standard evaluation, parameters passed into the function need to be specifically told when to be evaluated. The quo(), enquo(), quo_name(), and !! functions/operator  make this work, but it took a while to figure out!

```{r}

average_replicates <- function(chl, replicates = c("a", "b"), var = chl) {
  # these functions tell dplyr that we don't want it to do it's handy NSE
  # for the parameters we're passing
  quo_var <- enquo(var)
  var_name <- as_label(quo_var)
  quo_rep <- enquo(replicates)
  
  # the !! and := let dplyr know exactly when and how to do it's NSE 'magic'
  chl %>% 
    group_by(cruise, cast, niskin, filter_size) %>%
    filter(replicate %in% !!quo_rep) %>%
    summarize(!!var_name := mean(!!quo_var))
}
```

Now use the function to reduce our replicates and clean up a bit.

```{r}
# use the above function
chl_avg <- average_replicates(chl)

# now only include WSW
chl_avg <- filter(chl_avg, filter_size == '>0')
head(chl_avg)
```

Now the two datasets need to be collated so that they align on cruise, cast, and niskin bottle number.

The column containing CTD-derived data is called fleco_afl and the one containing the sample-derived data is called chl.

```{r}
merged <- inner_join(chl_avg, btl, by = c("cruise", "cast", "niskin"))
head(merged)
select(merged, cruise, cast, niskin, chl, fleco_afl, par) %>%
  head()
```

Plot it up!

```{r}
ggplot(merged, aes(chl, fleco_afl, fill = par)) +
  geom_point(shape = 21) +
  scale_color_viridis_c(aesthetics = "fill") +
  theme_bw() +
  labs(title = "CHL concentration: CTD vs. sampled",
       x = "sampled",
       y = "CTD")
```

